---
title: "03_modeling"
# output: html_document
---

# Environment

```{r}
library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(caret)

extract_coef_pval_OR <- function(model) {
  
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_CI_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_CI_upper = round(exp(estimate + 1.96 * std.error),2)
  )
  )
} 
```

# Load File

```{r}
final_dataset <- (
  readRDS(
    "~/projects/vacunas_brands/data/estudio_vacunasbrands_t_final_selected_manuscript.rds"
  )
)
```

# Analysis of variable group_vaccines

## Logistic Regression of Hospital Admission

### Fitting the model

```{r}
print('Logistic Regression of Hospital Admission')
print(Sys.time())
logreg_hadm <- glm(
   hosp_stay_bin ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + group_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_hadm_1 <- extract_coef_pval_OR(logreg_hadm)

write.csv(coef_pval_df_hadm_1,'~/projects/vacunas_brands/data/oddsratio_hadm.csv')
htmlTable(coef_pval_df_hadm_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for Hospital Admission")
```

### Forest Plot

```{r}
variable_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)",
  "Last vaccination/infection (<1 month ago)", "Last vaccination/infection (1-3 months ago)",
  "Last vaccination/infection (3-6 months ago)", "Last vaccination/infection (>6 months ago)", "Asthma",
  "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)",
  "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease",
  "HIV+", "Haematologic cancer", "Solid organ cancer", "Pfizer/BioNTech vaccine",
  "Moderna vaccine", "Mix of RNA vaccines", "AstraZeneca vaccine", "Janssen vaccine",
  "Mix of EU non-RNA vaccines", "Mix of EU RNA and non-RNA vaccines", "Sputnik V vaccine",
  "Sinopharm vaccine", "CoronaVac (Sinovac Biotech) vaccine", "Covishield vaccine", "Abdala vaccine",
  "Mix of EU RNA vaccines + non-EU Vaccines", "Mix of EU non-RNA Vaccines + non-EU Vaccines"
))

coef_pval_df_hadm_2 <- coef_pval_df_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "group_vaccinesPfizer/BioNTech" ~ "Pfizer/BioNTech vaccine",
      term == "group_vaccinesModerna" ~ "Moderna vaccine",
      term == "group_vaccinesMix of EU RNA vaccines" ~ "Mix of RNA vaccines",
      term == "group_vaccinesAstraZeneca" ~ "AstraZeneca vaccine",
      term == "group_vaccinesJanssen" ~ "Janssen vaccine",
      term == "group_vaccinesMix of EU non-RNA vaccines" ~ "Mix of EU non-RNA vaccines",
      term == "group_vaccinesMix of EU RNA and non-RNA vaccines" ~ "Mix of EU RNA and non-RNA vaccines",
      term == "group_vaccinesSputnik V" ~ "Sputnik V vaccine",
      term == "group_vaccinesSinopharm" ~ "Sinopharm vaccine",
      term == "group_vaccinesCoronaVac (Sinovac Biotech)" ~ "CoronaVac (Sinovac Biotech) vaccine",
      term == "group_vaccinesCovishield" ~ "Covishield vaccine",
      term == "group_vaccinesAbdala" ~ "Abdala vaccine",
      term == "group_vaccinesEU RNA vaccines + non-EU Vaccines" ~ "Mix of EU RNA vaccines + non-EU Vaccines",
      term == "group_vaccinesEU non-RNA Vaccines + non-EU Vaccines" ~ "Mix of EU non-RNA Vaccines + non-EU Vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order)
  ) %>%
  filter(term != "Intercept")


ggplot(coef_pval_df_hadm_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15, position = position_dodge(width = 0.5)) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Hospital admission') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))



# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_hadm.png", width = 6, height = 4, dpi = 300)
```

## Logistic Regression of ICU Admission

### Fitting the model

```{r}
print('Logistic Regression of Hospital Admission')
print(Sys.time())
logreg_icuadm <- glm(
   icu_stay_bin ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + group_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_icuadm_1 <- extract_coef_pval_OR(logreg_icuadm)

write.csv(coef_pval_df_icuadm_1,'~/projects/vacunas_brands/data/oddsratio_icuadm.csv')
htmlTable(coef_pval_df_icuadm_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for ICU Admission")
```

### Forest Plot

```{r}
coef_pval_df_icuadm_2<- coef_pval_df_icuadm_1  %>% 
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "group_vaccinesPfizer/BioNTech" ~ "Pfizer/BioNTech vaccine",
      term == "group_vaccinesModerna" ~ "Moderna vaccine",
      term == "group_vaccinesMix of EU RNA vaccines" ~ "Mix of RNA vaccines",
      term == "group_vaccinesAstraZeneca" ~ "AstraZeneca vaccine",
      term == "group_vaccinesJanssen" ~ "Janssen vaccine",
      term == "group_vaccinesMix of EU non-RNA vaccines" ~ "Mix of EU non-RNA vaccines",
      term == "group_vaccinesMix of EU RNA and non-RNA vaccines" ~ "Mix of EU RNA and non-RNA vaccines",
      term == "group_vaccinesSputnik V" ~ "Sputnik V vaccine",
      term == "group_vaccinesSinopharm" ~ "Sinopharm vaccine",
      term == "group_vaccinesCoronaVac (Sinovac Biotech)" ~ "CoronaVac (Sinovac Biotech) vaccine",
      term == "group_vaccinesCovishield" ~ "Covishield vaccine",
      term == "group_vaccinesAbdala" ~ "Abdala vaccine",
      term == "group_vaccinesEU RNA vaccines + non-EU Vaccines" ~ "Mix of EU RNA vaccines + non-EU Vaccines",
      term == "group_vaccinesEU non-RNA Vaccines + non-EU Vaccines" ~ "Mix of EU non-RNA Vaccines + non-EU Vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order)
  ) %>%
  filter(term != "Intercept")

ggplot(coef_pval_df_icuadm_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('ICU admission') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_icuadm.png", width = 6, height = 4, dpi = 300)
```

## Logistic Regression of Mortality at 30 days by COVID-19

### Fitting the model

```{r}
print('Logistic Regression of Mortality at 30 days by COVID-19')
print(Sys.time())
logreg_death <- glm(
   mortality30d ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + group_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_death_1 <- extract_coef_pval_OR(logreg_death)

write.csv(coef_pval_df_death_1,'~/projects/vacunas_brands/data/oddsratio_death.csv')
htmlTable(coef_pval_df_death_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for Mortality at 30 days by COVID-19")
```

### Forest Plot

```{r}
coef_pval_df_death_2<- coef_pval_df_death_1  %>% 
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "group_vaccinesPfizer/BioNTech" ~ "Pfizer/BioNTech vaccine",
      term == "group_vaccinesModerna" ~ "Moderna vaccine",
      term == "group_vaccinesMix of EU RNA vaccines" ~ "Mix of RNA vaccines",
      term == "group_vaccinesAstraZeneca" ~ "AstraZeneca vaccine",
      term == "group_vaccinesJanssen" ~ "Janssen vaccine",
      term == "group_vaccinesMix of EU non-RNA vaccines" ~ "Mix of EU non-RNA vaccines",
      term == "group_vaccinesMix of EU RNA and non-RNA vaccines" ~ "Mix of EU RNA and non-RNA vaccines",
      term == "group_vaccinesSputnik V" ~ "Sputnik V vaccine",
      term == "group_vaccinesSinopharm" ~ "Sinopharm vaccine",
      term == "group_vaccinesCoronaVac (Sinovac Biotech)" ~ "CoronaVac (Sinovac Biotech) vaccine",
      term == "group_vaccinesCovishield" ~ "Covishield vaccine",
      term == "group_vaccinesAbdala" ~ "Abdala vaccine",
      term == "group_vaccinesEU RNA vaccines + non-EU Vaccines" ~ "Mix of EU RNA vaccines + non-EU Vaccines",
      term == "group_vaccinesEU non-RNA Vaccines + non-EU Vaccines" ~ "Mix of EU non-RNA Vaccines + non-EU Vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order)
  ) %>%
  filter(term != "Intercept")

ggplot(coef_pval_df_death_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 30 days by COVID-19') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_death.png", width = 6, height = 4, dpi = 300)
```

# Analysis of variable location_vaccines

## Logistic Regression of Hospital Admission

### Fitting the model

```{r}
print('Logistic Regression of Hospital Admission')
print(Sys.time())
logreg_loc_hadm <- glm(
   hosp_stay_bin ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + european_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_loc_hadm_1 <- extract_coef_pval_OR(logreg_loc_hadm)

write.csv(coef_pval_df_loc_hadm_1,'~/projects/vacunas_brands/data/oddsratio_loc_hadm.csv')
htmlTable(coef_pval_df_loc_hadm_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for Hospital Admission")
```

### Forest Plot

```{r}
variable_loc_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)",
  "Last vaccination/infection (<1 month ago)", "Last vaccination/infection (1-3 months ago)",
  "Last vaccination/infection (3-6 months ago)", "Last vaccination/infection (>6 months ago)", "Asthma",
  "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)", "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease",
  "HIV+", "Haematologic cancer", "Solid organ cancer", "EU vaccines", "non-EU vaccines", "Mix of EU and non-EU vaccines"
))

coef_pval_df_loc_hadm_2 <- coef_pval_df_loc_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "european_vaccinesEU Vaccines" ~ "EU vaccines",
      term == "european_vaccinesNon-EU Vaccines" ~ "non-EU vaccines",
      term == "european_vaccinesEU and non-EU Vaccines" ~ "Mix of EU and non-EU vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_loc_order)
  ) %>%
  filter(term != "Intercept")


ggplot(coef_pval_df_loc_hadm_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15, position = position_dodge(width = 0.5)) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Hospital admission') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))



# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_loc_hadm.png", width = 6, height = 4, dpi = 300)
```

## Logistic Regression of ICU Admission

### Fitting the model

```{r}
print('Logistic Regression of Hospital Admission')
print(Sys.time())
logreg_loc_icuadm <- glm(
   icu_stay_bin ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + european_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_loc_icuadm_1 <- extract_coef_pval_OR(logreg_loc_icuadm)

write.csv(coef_pval_df_loc_icuadm_1,'~/projects/vacunas_brands/data/oddsratio_loc_icuadm.csv')
htmlTable(coef_pval_df_loc_icuadm_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for ICU Admission")
```

### Forest Plot

```{r}
coef_pval_df_loc_icuadm_2<- coef_pval_df_loc_icuadm_1  %>% 
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "european_vaccinesEU Vaccines" ~ "EU vaccines",
      term == "european_vaccinesNon-EU Vaccines" ~ "non-EU vaccines",
      term == "european_vaccinesEU and non-EU Vaccines" ~ "Mix of EU and non-EU vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_loc_order)
  ) %>%
  filter(term != "Intercept")

ggplot(coef_pval_df_loc_icuadm_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('ICU admission') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_loc_icuadm.png", width = 6, height = 4, dpi = 300)
```

## Logistic Regression of Mortality at 30 days by COVID-19

### Fitting the model

```{r}
print('Logistic Regression of Mortality at 30 days by COVID-19')
print(Sys.time())
logreg_loc_death <- glm(
   mortality30d ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + ultima_inmun_infecc_bin + european_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_loc_death_1 <- extract_coef_pval_OR(logreg_loc_death)

write.csv(coef_pval_df_loc_death_1,'~/projects/vacunas_brands/data/oddsratio_loc_death.csv')
htmlTable(coef_pval_df_loc_death_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for Mortality at 30 days by COVID-19")
```

### Forest Plot

```{r}
coef_pval_df_loc_death_2<- coef_pval_df_loc_death_1  %>% 
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ / Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "ultima_inmun_infecc_binLess than one month" ~ "Last vaccination/infection (<1 month ago)",
      term == "ultima_inmun_infecc_binOne to three months" ~ "Last vaccination/infection (1-3 months ago)",
      term == "ultima_inmun_infecc_binThree to six months" ~ "Last vaccination/infection (3-6 months ago)",
      term == "ultima_inmun_infecc_binSix months or more" ~ "Last vaccination/infection (>6 months ago)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "european_vaccinesEU Vaccines" ~ "EU vaccines",
      term == "european_vaccinesNon-EU Vaccines" ~ "non-EU vaccines",
      term == "european_vaccinesEU and non-EU Vaccines" ~ "Mix of EU and non-EU vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_loc_order)
  ) %>%
  filter(term != "Intercept")

ggplot(coef_pval_df_loc_death_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 30 days by COVID-19') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/fp_loc_death.png", width = 6, height = 4, dpi = 300)
```

