---
title: "03_modeling_clinicalevent"
# output: html_document
---

# Environment

```{r}
library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(caret)
library(kableExtra)

extract_coef_pval_OR <- function(model) {
  
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_CI_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_CI_upper = round(exp(estimate + 1.96 * std.error),2)
  )
  )
} 
```

# Load File

```{r}
final_dataset <- (
  readRDS(
    "~/projects/vacunas_brands/Rfiles/estudio_vacunasbrands_t_final_selected_manuscript.rds"
  )
)
```

# Analysis of variable eu_vaccines

## Logistic Regression of Clinical Event

### Fitting the model

```{r}
print('Logistic Regression of Clinical Event')
print(Sys.time())
logreg_hadm <- glm(
   clinical_event ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + eu_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_hadm_1 <- extract_coef_pval_OR(logreg_hadm)

```

### Forest Plot

```{r}
variable_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Asthma", "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)", "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease", "HIV+", "Haematologic cancer", "Solid organ cancer", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)", "Moderna vaccine", "Pfizer/BioNTech vaccine", "AstraZeneca vaccine", "Janssen vaccine"))

coef_pval_df_hadm_2 <- coef_pval_df_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ/Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "eu_vaccinesModerna" ~ "Moderna vaccine",
      term == "eu_vaccinesPfizer/BioNTech" ~ "Pfizer/BioNTech vaccine",
      term == "eu_vaccinesAstraZeneca" ~ "AstraZeneca vaccine",
      term == "eu_vaccinesJanssen" ~ "Janssen vaccine",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order), 
    adverse_protective = case_when(
           OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
           OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
           TRUE ~ "Inconclusive"
    ),
    adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive"))) %>%
  filter(term != "Intercept")

# Save OR table and print in console
write.csv(coef_pval_df_hadm_2,'~/projects/vacunas_brands/data/oddsratio_eu_vaccines_clinicalevent.csv')
htmlTable(coef_pval_df_hadm_2[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "EU-approved Vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")

# Export OR table as kable 
or_table <- coef_pval_df_hadm_2 %>%
  select(term, OR, OR_CI_lower, OR_CI_upper) %>%
  mutate(
    term = factor(term, levels = variable_order),
    OR_CI_upper = ifelse(OR_CI_upper <= 100, sprintf("%.2f", OR_CI_upper), sprintf("%.2e", OR_CI_upper))) %>%
  arrange(desc(term)) %>%
  kbl(caption = "EU-approved Vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  save_kable("~/projects/vacunas_brands/figures/OR/OR_euvaccines_clinicalevent.jpg",
             zoom = 10)

# Build the Forest plot
ggplot(coef_pval_df_hadm_2, 
       aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "Logistic Regression. Odds Ratio and Confidence Intervals \nfor Composite Outcome-associated Factors. \nEU-approved Vaccines") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome") +
  scale_x_continuous(limits = c(0, 5)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))

ggsave("~/projects/vacunas_brands/figures/Forest_plot/fp_euvaccines_clinicalevent.png", width = 8, height = 6, dpi = 300)

```

# Analysis of variable eu_and_noneu_vaccines

## Logistic Regression of Clinical Event

### Fitting the model

```{r}
print('Logistic Regression of Clinical Event')
print(Sys.time())
logreg_hadm <- glm(
   clinical_event ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + eu_and_noneu_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_hadm_1 <- extract_coef_pval_OR(logreg_hadm)

```

### Forest Plot

```{r}
variable_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Asthma", "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)", "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease", "HIV+", "Haematologic cancer", "Solid organ cancer", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)", "EU-approved Vaccines", "Non EU-approved Vaccines", "Mix of EU-approved and non EU-approved Vaccines"))

coef_pval_df_hadm_2 <- coef_pval_df_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ/Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "eu_and_noneu_vaccinesEU Vaccines" ~ "EU-approved Vaccines",
      term == "eu_and_noneu_vaccinesNon-EU Vaccines" ~ "Non EU-approved Vaccines",
      term == "eu_and_noneu_vaccinesEU and non-EU Vaccines" ~ "Mix of EU-approved and non EU-approved Vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order), 
    adverse_protective = case_when(
           OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
           OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
           TRUE ~ "Inconclusive"
    ),
    adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive"))) %>%
  filter(term != "Intercept")

# Save OR table and print in console
write.csv(coef_pval_df_hadm_2,'~/projects/vacunas_brands/data/oddsratio_eu_and_noneu_vaccines_clinicalevent.csv')
htmlTable(coef_pval_df_hadm_2[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "EU vs non-EU approved Vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")

# Export OR table as kable 
or_table <- coef_pval_df_hadm_2 %>%
  select(term, OR, OR_CI_lower, OR_CI_upper) %>%
  mutate(
    term = factor(term, levels = variable_order),
    OR_CI_upper = ifelse(OR_CI_upper <= 100, sprintf("%.2f", OR_CI_upper), sprintf("%.2e", OR_CI_upper))) %>%
  arrange(desc(term)) %>%
  kbl(caption = "EU-approved and non EU-approved vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  save_kable("~/projects/vacunas_brands/figures/OR/OR_eu-and-non-euvaccines_clinicalevent.jpg", zoom = 10)

# Build the Forest plot
ggplot(coef_pval_df_hadm_2, 
       aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "Logistic Regression. Odds Ratio and Confidence Intervals \nfor Composite Outcome-associated Factors. \nEU vs non-EU approved Vaccines") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome") +
  scale_x_continuous(limits = c(0, 5)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))

# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/Forest_plot/fp_euandnoneu_vaccines_clinicalevent.png", width = 8, height = 6, dpi = 300)
```

# Analysis of variable rna_vaccines

## Logistic Regression of Clinical Event

### Fitting the model

```{r}
print('Logistic Regression of Clinical Event')
print(Sys.time())
logreg_hadm <- glm(
    clinical_event ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + rna_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_hadm_1 <- extract_coef_pval_OR(logreg_hadm)

```

### Forest Plot

```{r}
variable_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Asthma", "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)", "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease", "HIV+", "Haematologic cancer", "Solid organ cancer", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)", "Mix of EU RNA vaccines", "Mix of EU non-RNA vaccines", "Mix of EU RNA and non-RNA vaccines"))

coef_pval_df_hadm_2 <- coef_pval_df_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ/Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "rna_vaccinesMix of EU RNA vaccines" ~ "Mix of EU RNA vaccines",
      term == "rna_vaccinesMix of EU non-RNA vaccines" ~ "Mix of EU non-RNA vaccines",
      term == "rna_vaccinesMix of EU RNA and non-RNA vaccines" ~ "Mix of EU RNA and non-RNA vaccines",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order), 
    adverse_protective = case_when(
           OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
           OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
           TRUE ~ "Inconclusive"
    ),
    adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive"))) %>%
  filter(term != "Intercept")

# Save OR table and print in console
write.csv(coef_pval_df_hadm_2,'~/projects/vacunas_brands/data/oddsratio_rnavaccines_clinicalevent.csv')
htmlTable(coef_pval_df_hadm_2[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "RNA vs non-RNA Vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")

# Export OR table as kable 
or_table <- coef_pval_df_hadm_2 %>%
  select(term, OR, OR_CI_lower, OR_CI_upper) %>%
  mutate(
    term = factor(term, levels = variable_order),
    OR_CI_upper = ifelse(OR_CI_upper <= 100, sprintf("%.2f", OR_CI_upper), sprintf("%.2e", OR_CI_upper))) %>%
  arrange(desc(term)) %>%
  kbl(caption = "RNA vs non-RNA Vaccines - Odds Ratio and Confidence Intervals for composite outcome-associated Factors" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  save_kable("~/projects/vacunas_brands/figures/OR/OR_rna-vaccines_clinicalevent.jpg",
             zoom = 10)

# Build the Forest plot
ggplot(coef_pval_df_hadm_2, 
       aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "Logistic Regression. Odds Ratio and Confidence Intervals \nfor Composite Outcome-associated Factors. \nRNA vs non-RNA Vaccines") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome") +
  scale_x_continuous(limits = c(0, 5)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))

# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/Forest_plot/fp_rnavaccines_clinicalevent.png", width = 8, height = 6, dpi = 300)
```
# Analysis of variable combined_vaccines

## Logistic Regression of Clinical Event

### Fitting the model

```{r}
print('Logistic Regression of Clinical Event')
print(Sys.time())
logreg_hadm <- glm(
    clinical_event ~ age + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc +  hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido + n_prev_infecc + combined_vaccines,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

### Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_hadm_1 <- extract_coef_pval_OR(logreg_hadm)
```

### Forest Plot

```{r}
variable_order <- rev(c(
  "Intercept", "Age", "Gender (male)", "Asthma", "Ischemic heart disease", "Hepatic cirrhosis", "Diabetes", "Chronic Obstructive Pulmonary Disease (COPD)", "Chronic Hepatic Disease (except cirrhosis)", "Hypertension", "Heart failure", "Chronic kidney disease", "HIV+", "Haematologic cancer", "Solid organ cancer", "Period of α predominance",
  "α-Δ transition period", "Period of Δ predominance", "Δ-Ό transition period",
  "Period of Ό predominance", "Number of Previous Infections (1)", "Number of Previous Infections (2)", "Pfizer/BioNTech+Pfizer/BioNTech", "Pfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech", "Moderna+Moderna", "AstraZeneca+AstraZeneca", "Pfizer/BioNTech", "Janssen", "AstraZeneca", "Moderna", "AstraZeneca+AstraZeneca+Moderna", "AstraZeneca+AstraZeneca+Pfizer/BioNTech", "Pfizer/BioNTech+Pfizer/BioNTech+Moderna", "Janssen+Pfizer/BioNTech", "Moderna+Moderna+Pfizer/BioNTech","Janssen+Moderna",  "AstraZeneca+Pfizer/BioNTech", "Pfizer/BioNTech+Moderna", "Moderna+Pfizer/BioNTech", "AstraZeneca+Pfizer/BioNTech+Pfizer/BioNTech"))

coef_pval_df_hadm_2 <- coef_pval_df_hadm_1  %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "age" ~ "Age",
      term == "desc_sexoHombre" ~ "Gender (male)",
      term == "variant_periodα" ~ "Period of α predominance",
      term == "variant_periodα/Δ transition" ~ "α-Δ transition period",
      term == "variant_periodΔ" ~ "Period of Δ predominance",
      term == "variant_periodΔ/Ό transition" ~ "Δ-Ό transition period",
      term == "variant_periodΌ" ~ "Period of Ό predominance",
      term == "n_prev_infeccUno" ~ "Number of Previous Infections (1)",
      term == "n_prev_infeccDos" ~ "Number of Previous Infections (2)",
      term == "asmaSí" ~ "Asthma",
      term == "cardiopatia_isquemicaSí" ~ "Ischemic heart disease",
      term == "cirrosis_hepaticaSí" ~ "Hepatic cirrhosis",
      term == "diabetesSí" ~ "Diabetes",
      term == "epocSí" ~ "Chronic Obstructive Pulmonary Disease (COPD)",
      term == "hepatopatia_cronica_excepto_cirrosisSí" ~ "Chronic Hepatic Disease (except cirrhosis)",
      term == "hipertensionSí" ~ "Hypertension",
      term == "insuficiencia_cardiacaSí" ~ "Heart failure",
      term == "insuficiencia_renal_cronicaSí" ~ "Chronic kidney disease",
      term == "vihSí" ~ "HIV+",
      term == "cancer_hematologicoSí" ~ "Haematologic cancer",
      term == "cancer_organo_solidoSí" ~ "Solid organ cancer",
      term == "combined_vaccinesPfizer/BioNTech+Pfizer/BioNTech" ~ "Pfizer/BioNTech+Pfizer/BioNTech",
      term == "combined_vaccinesPfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech" ~ "Pfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech",
      term == "combined_vaccinesModerna+Moderna" ~ "Moderna+Moderna",
      term == "combined_vaccinesAstraZeneca+AstraZeneca" ~ "AstraZeneca+AstraZeneca",
      term == "combined_vaccinesPfizer/BioNTech" ~ "Pfizer/BioNTech",
      term == "combined_vaccinesJanssen" ~ "Janssen",
      term == "combined_vaccinesAstraZeneca" ~ "AstraZeneca",
      term == "combined_vaccinesModerna" ~ "Moderna",
      term == "combined_vaccinesJanssen+Moderna" ~ "Janssen+Moderna",
      term == "combined_vaccinesPfizer/BioNTech+Pfizer/BioNTech+Moderna" ~ "Pfizer/BioNTech+Pfizer/BioNTech+Moderna",
      term == "combined_vaccinesAstraZeneca+Pfizer/BioNTech" ~ "AstraZeneca+Pfizer/BioNTech",
      term == "combined_vaccinesJanssen+Pfizer/BioNTech" ~ "Janssen+Pfizer/BioNTech",
      term == "combined_vaccinesAstraZeneca+AstraZeneca+Pfizer/BioNTech" ~ "AstraZeneca+AstraZeneca+Pfizer/BioNTech",
      term == "combined_vaccinesModerna+Moderna+Pfizer/BioNTech" ~ "Moderna+Moderna+Pfizer/BioNTech",
      term == "combined_vaccinesAstraZeneca+AstraZeneca+Moderna" ~ "AstraZeneca+AstraZeneca+Moderna",
      term == "combined_vaccinesModerna+Pfizer/BioNTech" ~ "Moderna+Pfizer/BioNTech",
      term == "combined_vaccinesAstraZeneca+Pfizer/BioNTech+Pfizer/BioNTech" ~ "AstraZeneca+Pfizer/BioNTech+Pfizer/BioNTech",
      term == "combined_vaccinesPfizer/BioNTech+Moderna" ~ "Pfizer/BioNTech+Moderna",
      TRUE ~ term
    ),
    term = factor(term, levels = variable_order), 
    adverse_protective = case_when(
           OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
           OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
           TRUE ~ "Inconclusive"
    ),
    adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive"))) %>%
  filter(term != "Intercept")

# Save OR table and print in console
write.csv(coef_pval_df_hadm_2,'~/projects/vacunas_brands/data/oddsratio_combinedvaccines_clinicalevent.csv')
htmlTable(coef_pval_df_hadm_2[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Vaccine Shot Combinations - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")

# Export OR table as kable 
or_table <- coef_pval_df_hadm_2 %>%
  select(term, OR, OR_CI_lower, OR_CI_upper) %>%
  mutate(
    term = factor(term, levels = variable_order),
    OR_CI_upper = ifelse(OR_CI_upper <= 100, sprintf("%.2f", OR_CI_upper), sprintf("%.2e", OR_CI_upper))) %>%
  arrange(desc(term)) %>%
  kbl(caption = "Vaccine shot combinations - Odds Ratio and Confidence Intervals for composite outcome-associated Factors" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria") %>%
  save_kable("~/projects/vacunas_brands/figures/OR/OR_combined-vaccines_clinicalevent.jpg",zoom = 10)

# Build the Forest plot
ggplot(coef_pval_df_hadm_2, 
       aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "Logistic Regression. Odds Ratio and Confidence Intervals\nfor Composite Outcome-associated Factors. \nVaccine Shot Combinations") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome") +
  scale_x_continuous(limits = c(0, 5)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.title.position = "plot",
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))

# Save the plot as a PNG file
ggsave("~/projects/vacunas_brands/figures/Forest_plot/fp_combinedvaccines_clinicalevent.png", width = 8, height = 8, dpi = 300)
```
