---
title: "03_modeling_clinicalevent"
# output: html_document
---

# Environment

```{r}
library(skimr)
library(mgcv)
# library(parallel)
# library(parglm)
# library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(caret)
library(kableExtra)
library(magrittr)
# library(plotly)
library(htmlwidgets)
# library(webshot2)
```

# Load File

```{r}
final_dataset <- (
  readRDS(
    "~/projects/vacunas_brands_2025/Rfiles/final_dataset_brands_excluded.rds"
  )
)

# final_dataset <- final_dataset[1:10000,]
```

# Vectors/Dictionaries
```{r}

dict_terms <- c(
  "age" = "Age",
  "age_grouped12-30" = "Age (12-30)",
  "age_grouped31-65" = "Age (31-65)",
  "age_grouped>65" = "Age (>65)",
  "ultima_inmun_infecc_groupedNone" = "No previous immunization",
  "ultima_inmun_infecc_grouped14-45" = "Last immunization: 14-45 days",
  "ultima_inmun_infecc_grouped46-90" = "Last immunization: 46-90 days",
  "ultima_inmun_infecc_grouped91-170" = "Last immunization: 91-170 days",
  "ultima_inmun_infecc_grouped171-247" = "Last immunization: 171-247 days",
  "ultima_inmun_infecc_grouped>247" = "Last immunization: >247 days",
  "desc_sexoHombre" = "Sex (male)",
  "variant_periodα" = "Period of α predominance",
  "variant_periodα/Δ transition" = "α-Δ transition period",
  "variant_periodΔ" = "Period of Δ predominance",
  "variant_periodΔ/Ό transition" = "Δ-Ό transition period",
  "variant_periodΌ" = "Period of Ό predominance",
  "asmaSí" = "Asthma",
  "cardiopatia_isquemicaSí" = "Ischemic heart disease",
  "cirrosis_hepaticaSí" = "Hepatic cirrhosis",
  "diabetesSí" = "Diabetes",
  "epocSí" = "Chronic Obstructive Pulmonary Disease (COPD)",
  "hepatopatia_cronica_excepto_cirrosisSí" = "Chronic Hepatic Disease (except cirrhosis)",
  "hipertensionSí" = "Hypertension",
  "insuficiencia_cardiacaSí" = "Heart failure",
  "insuficiencia_renal_cronicaSí" = "Chronic kidney disease",
  "vihSí" = "HIV+",
  "cancer_hematologicoSí" = "Haematologic cancer",
  "cancer_organo_solidoSí" = "Solid organ cancer",
  "ultima_inmun_infecc_binNone" = "No vaccination/infection",
  "ultima_inmun_infecc_bin31-60" = "Last vaccination/infection (31-60 days ago)",
  "ultima_inmun_infecc_bin61-90" = "Last vaccination/infection (61-90 days ago)",
  "ultima_inmun_infecc_bin91-120" = "Last vaccination/infection (91-120 days ago)",
  "ultima_inmun_infecc_bin121-180" = "Last vaccination/infection (121-180 days ago)",
  "ultima_inmun_infecc_bin>180" = "Last vaccination/infection (>180 days ago)",
  "vac_1shotAstraZeneca" = "AstraZeneca",
  "vac_1shotJanssen" = "Janssen",
  "vac_1shotModerna" = "Moderna",
  "vac_1shotPfizer/BioNTech" = "Pfizer/BioNTech",
  "vac_1shotOther" = "Other",
  "vac_2shotAstraZeneca+AstraZeneca" = "AstraZeneca+AstraZeneca",
  "vac_2shotModerna+Moderna" = "Moderna+Moderna",
  "vac_2shotPfizer/BioNTech+Pfizer/BioNTech" = "Pfizer/BioNTech+Pfizer/BioNTech",
  "vac_2shotOther" = "Other",
  "vac_3shotAstraZeneca+AstraZeneca+Moderna" = "AstraZeneca+AstraZeneca+Moderna",
  "vac_3shotAstraZeneca+AstraZeneca+Pfizer/BioNTech" = "AstraZeneca+AstraZeneca+Pfizer/BioNTech",
  "vac_3shotModerna+Moderna+Moderna" = "Moderna+Moderna+Moderna",
  "vac_3shotPfizer/BioNTech+Pfizer/BioNTech+Moderna" = "Pfizer/BioNTech+Pfizer/BioNTech+Moderna",
  "vac_3shotPfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech" = "Pfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech"
)

  # Variable mapping and ordering
variable_order_reduced <- rev(c(
"Intercept", "Age", 
"Age (12-30)", "Age (31-65)", "Age (>65)", 
"AstraZeneca", "Janssen", "Moderna", "Pfizer/BioNTech",
"AstraZeneca+AstraZeneca", "Moderna+Moderna", "Pfizer/BioNTech+Pfizer/BioNTech",
"AstraZeneca+AstraZeneca+Moderna", "AstraZeneca+AstraZeneca+Pfizer/BioNTech", "Moderna+Moderna+Moderna",
"Pfizer/BioNTech+Pfizer/BioNTech+Moderna", "Pfizer/BioNTech+Pfizer/BioNTech+Pfizer/BioNTech", 
"Last immunization: 14-45 days", "No previous immunization", "Last immunization: 46-90 days", "Last immunization: 91-170 days", "Last immunization: 171-247 days", "Last immunization: >247 days"
  ))

added_variables <- rev(c(
  "Sex (male)",
  "Period of α predominance",
  "α-Δ transition period",
  "Period of Δ predominance",
  "Δ-Ό transition period",
  "Period of Ό predominance",
  "Asthma",
  "Ischemic heart disease",
  "Hepatic cirrhosis",
  "Diabetes",
  "Chronic Obstructive Pulmonary Disease (COPD)",
  "Chronic Hepatic Disease (except cirrhosis)",
  "Hypertension",
  "Heart failure",
  "Chronic kidney disease",
  "HIV+",
  "Haematologic cancer",
  "Solid organ cancer"
))

variable_order_all <- c(added_variables, variable_order_reduced)
  
# Independent variables (without vaccine variable)
predictor_vars <- c("age", "ultima_inmun_infecc_dias", "desc_sexo", "variant_period", "asma", "cardiopatia_isquemica", "cirrosis_hepatica", "diabetes", "epoc", "hepatopatia_cronica_excepto_cirrosis", "hipertension", "insuficiencia_cardiaca", "insuficiencia_renal_cronica", "vih", "cancer_hematologico", "cancer_organo_solido")
# Independent variables (spline)
spline_vars <- c("age")

```

# Functions

## Extract OR
```{r}
# Extracts and formats logistic regression model coefficients into a data frame with OR and CI
extract_coef_pval_OR <- function(model) {
  
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_CI_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_CI_upper = round(exp(estimate + 1.96 * std.error),2)
  )
  )
} 

```

## GAM model with 1 spline (age)
```{r}

# Function for GAM-based logistic analysis and forest plot
log_reg_gam_1s <- function(data, vaccine_variable, variables, file_name, title) {
  
  print('GAM (logistic link) analysis of Clinical Event')
  start_time <- Sys.time() 
  print(Sys.time())
  
  # Define the order of variables for the model based on the 'variables' argument
  if (variables == "reduced") {
    variable_order <- variable_order_reduced
    caption_confounders <- "Comorbidities, sex and SARS-CoV-2 variant period were also included as confounders"
  } else if (variables == "all") {
    variable_order <- variable_order_all
    caption_confounders <- ""
  } else {
    # Handle invalid values for 'variables'
    stop("The 'variables' argument must be 'reduced' or 'all'")
  }
  
# Build formula dynamically
  formula_text <- paste0( 
    "clinical_event ~ s(age) + ultima_inmun_infecc_grouped + ", vaccine_variable, " + desc_sexo + variant_period + asma + cardiopatia_isquemica + cirrosis_hepatica + diabetes + epoc + hepatopatia_cronica_excepto_cirrosis + hipertension + insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido"
    )
formula_glm <- as.formula(formula_text)

  ## Fit GAM model (logistic link)
  model <- gam(
  formula = formula_glm,
  data = data,
  family = binomial(), 
  method = "REML"
)
  
  print(summary(model))
  
  ## Extract odds ratios
  model_df_1 <- extract_coef_pval_OR(model)
  
  model_df_2 <- model_df_1 %>%
    mutate(
      term = case_when(
        term == "(Intercept)" ~ "Intercept",
        TRUE ~ coalesce(dict_terms[term], term)  # Use dictionary mapping
      ),
      term = factor(term, levels = variable_order),
      adverse_protective = case_when(
        OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
        OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
        TRUE ~ "Inconclusive"
      ),
      adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive")), 
    ) %>%
    filter(term != "Intercept")
  
  ## Save OR table
  write.csv(
    model_df_2 %>%
      arrange(factor(term, levels = rev(variable_order))),
    paste0("~/projects/vacunas_brands_2025/data/OR_", file_name, ".csv"),
    row.names = FALSE
  )
  
  ## Generate HTML table
  html_output <- htmlTable(
    model_df_2 %>%
      select(term, OR, OR_CI_lower, OR_CI_upper, adverse_protective) %>%
      mutate(term = factor(term, levels = variable_order)) %>%
      arrange(factor(term, levels = rev(variable_order))),
  caption = paste0(title, " - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")
)
  print(html_output)
  
  ## Create the plot with ggplot2
  plot <- ggplot(model_df_2, 
                 aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term, color = adverse_protective)) +
    ### Anchor legend with all categories
    geom_point(                                            
      data = data.frame(
        OR = 1, OR_CI_lower = 1, OR_CI_upper = 1,
        term = factor(rep(levels(model_df_2$term)[1], 3), levels = levels(model_df_2$term)),
        adverse_protective = factor(c("Adverse","Protective","Inconclusive"),
                                    levels = c("Adverse","Protective","Inconclusive"))
      ),
      mapping = aes(x = OR, y = term, color = adverse_protective),
      inherit.aes = FALSE,
      alpha = 0 ##invisible points to ensure all legend categories appear    
    ) +
    geom_errorbarh(height = 0.2, na.rm = TRUE, show.legend = TRUE) +
    geom_point(size = 2, na.rm = TRUE, shape = 16, show.legend = TRUE) +       
    geom_vline(xintercept = 1, linetype = "dashed") +
    labs(x = "Odds Ratio", y = NULL, 
         title = paste0("GAM (spline for age)\n Odds Ratios for Composite Outcome-associated Factors\n", title), 
         caption = caption_confounders
         ) +
    theme_minimal() +
    scale_color_manual(
      values = c("Adverse" = "#c0392b", "Protective" = "#27ae60", "Inconclusive" = "gray70"),
      breaks = c("Adverse", "Protective", "Inconclusive"),
      limits = c("Adverse", "Protective", "Inconclusive"),
      drop = FALSE, ##show all legend categories even if no data exists for some categories
      name = "Effect on Outcome") +
    scale_x_continuous(limits = c(0, 5)) +
    theme(
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      plot.title = element_text(hjust = 0.5),
      plot.title.position = "plot",
      legend.position = "top",
      axis.ticks = element_blank(),
      axis.text.x = element_text(color = "#2c3e50"),
      axis.text.y = element_text(color = "#2c3e50"), 
      ## Center bottom caption to the center
      plot.caption.position = "plot", plot.caption = element_text(
        hjust = 0.5,
        face = "italic",
        size = 9,
        margin = margin(t = 6)
      ))
  
  print(plot)

  ## Save plot
  ggsave(
    filename = paste0("~/projects/vacunas_brands_2025/figures/Forest_plot/fp_gam_1spline_", file_name, ".png"),
    plot = plot, 
    width = 8, 
    height = 6, 
    dpi = 300
  )
  
  time_elapsed <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  print(paste("Time to run:", time_elapsed, "minutes"))
}


```


## GAM model with 2 splines (not used)
```{r}

# Function for Logistic Regression Analysis
log_reg_gam_2s <- function(data, vaccine_variable, file_name, title, number_splines) {
  
  print('Logistic Regression of Clinical Event')
  start_time <- Sys.time() 
  print(Sys.time())
  
  ## Change independent variables for spline of age + last immunization days
  if (number_splines == 2) {
    # Add 'ultima_inmun_infecc_dias' to spline_vars
    spline_vars <- c("age", "ultima_inmun_infecc_dias")
    # And remove 'ultima_inmun_infecc_bin' from predictor_vars
    predictor_vars <- predictor_vars[-which(predictor_vars == "ultima_inmun_infecc_bin")]
  }
  
  ## Define spline terms
  spline_terms <- paste0("s(", spline_vars, ")")
  # Combine predictor variables with spline terms
  all_predictors <- c(spline_terms, predictor_vars)
  # Create formula
  formula <- as.formula(paste("clinical_event", "~", paste(all_predictors, collapse = " + "), " + ", vaccine_variable))
  
  ## Fit logistic regression model
  model <- gam(
    formula,
    data = data,
    family = binomial(),
    method = "REML"
  )
  
  print(summary(model))
  
  ## Extract odds ratios
  model_df_1 <- extract_coef_pval_OR(model)
  
  model_df_2 <- model_df_1 %>%
    mutate(
      term = case_when(
        term == "(Intercept)" ~ "Intercept",
        TRUE ~ coalesce(dict_terms[term], term)  # Use dictionary mapping
      ),
      term = factor(term, levels = variable_order),
      adverse_protective = case_when(
        OR_CI_lower > 1 & OR_CI_upper > 1 ~ "Adverse",
        OR_CI_lower < 1 & OR_CI_upper < 1 ~ "Protective",
        TRUE ~ "Inconclusive"
      ),
      adverse_protective = factor(adverse_protective, levels = c("Adverse", "Protective", "Inconclusive"))
    ) %>%
    filter(term != "Intercept")
  
  ## Save OR table
  write.csv(
    model_df_2 %>%
      arrange(factor(term, levels = rev(variable_order))),
    paste0("~/projects/vacunas_brands_2025/data/OR_", file_name, ".csv"),
    row.names = FALSE
  )
  
  ## Generate HTML table
  html_output <- htmlTable(
    model_df_2 %>%
      select(term, OR, OR_CI_lower, OR_CI_upper, adverse_protective) %>%
      mutate(term = factor(term, levels = variable_order)) %>%
      arrange(factor(term, levels = rev(variable_order))),
  caption = paste0(title, " - Odds Ratio and Confidence Intervals for composite outcome-associated Factors")
)
  print(html_output)
  
  ## Create the plot with ggplot2
  plot <- ggplot(model_df_2, 
                 aes(x = OR, xmin = OR_CI_lower, xmax = OR_CI_upper, y = term, color = adverse_protective)) +
    geom_point(size = 2, na.rm = TRUE) +  
    geom_errorbarh(height = 0.2, na.rm = TRUE) +
    geom_vline(xintercept = 1, linetype = "dashed") +
    labs(x = "Odds Ratio", y = NULL, title = paste0("Logistic Regression. Odds Ratio and Confidence Intervals \nfor Composite Outcome-associated Factors.\n", title)) +
    theme_minimal() +
    scale_color_manual(
      values = c("Adverse" = "#c0392b", "Protective" = "#27ae60", "Inconclusive" = "gray70"),
      limits = c("Adverse", "Protective", "Inconclusive"), #ensure all levels appear in legend
      labels = c("Adverse", "Protective", "Inconclusive"),
      name = "Effect on Outcome",
      guide = guide_legend(override.aes = list(size = 3)) 
    ) +
    scale_x_continuous(limits = c(0, 5)) +
    theme(
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      plot.title = element_text(hjust = 0.5),
      plot.title.position = "plot",
      legend.position = "top",
      axis.ticks = element_blank(),
      axis.text.x = element_text(color = "#2c3e50"),
      axis.text.y = element_text(color = "#2c3e50"))
  
  print(plot)

  ## Save plot
  ggsave(
    filename = paste0("~/projects/vacunas_brands_2025/figures/Forest_plot/fp_gam_2spline_", file_name, ".png"),
    plot = plot, 
    width = 8, 
    height = 6, 
    dpi = 300
  )
  
   time_elapsed <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  print(paste("Time to run:", time_elapsed, "minutes"))
}

```


## GAM plot

```{r}
# Definir la función para ajustar el modelo GAM y crear el gráfico
generate_gam_plot <- function(data, vaccine_variable, spline_var) {
  # Ajuste del modelo GAM
  print('GAM linear regression of Clinical Event')
  start_time <- Sys.time()
  print(Sys.time())

  # Definir el orden dinámico de las smooths
  if (spline_var == "age") {
    order_smooth <- "s(age) + s(ultima_inmun_infecc_dias)"
  } else if (spline_var == "ultima_inmun_infecc_dias") {
    order_smooth <- "s(ultima_inmun_infecc_dias) + s(age)"
  } else {
    stop("spline_var debe ser 'age' o 'ultima_inmun_infecc_dias'")
  }

  # Crear la fórmula completa como texto
  formula_text <- paste0(
    "clinical_event ~ ",
    order_smooth, " + ",
    vaccine_variable, " + desc_sexo + variant_period + asma + cardiopatia_isquemica + ",
    "cirrosis_hepatica + diabetes + epoc + hepatopatia_cronica_excepto_cirrosis + hipertension + ",
    "insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido"
  )

  # Convertir a fórmula
  formula_gam <- as.formula(formula_text)

  # Ajustar modelo
  model <- gam(
    formula = formula_gam,
    data = data,
    family = binomial(),
    method = "REML"
  )

  print(summary(model))

  # # Imprimir los nudos para las variables spline
  # print("Nudos para 'ultima_inmun_infecc_dias':")
  # print(model$smooth[[1]]$xp)  # Puntos de predicción (equivalente a nudos)
  #
  # print("Knots si están definidos explícitamente:")
  # print(model$smooth[[1]]$knots)
  #
  # print("Nombre de la variable spline:")
  # print(model$smooth[[1]]$term)


  # Función para extraer los datos del modelo GAM y calcular probabilidades
  extract_gam_data <- function(model, spline_var) {
    gam_plot_data <- plot(model, term = spline_var, pages = 1)[[1]] # Use term argument
    x <- gam_plot_data$x
    fit <- gam_plot_data$fit
    se <- gam_plot_data$se

    prob <- exp(fit) / (1 + exp(fit)) * 100
    lower <- exp(fit - 2 * se) / (1 + exp(fit - 2 * se)) * 100
    upper <- exp(fit + 2 * se) / (1 + exp(fit + 2 * se)) * 100

    data.frame(spline_var = x, probability = prob, lower = lower, upper = upper)
  }

# Función para crear el gráfico GAM
  create_gam_plot <- function(data, title, spline_var) {
    ggplot(data, aes(x = spline_var, y = probability)) +
      geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3) +
      geom_line() +
      labs(title = title, x = spline_var, y = "Probability (%)") +
      ylim(0, 100) +
      theme_minimal()
  }

  # Generar los datos para el gráfico y crear el gráfico
  plot_data <- extract_gam_data(model, spline_var)
  plot <- create_gam_plot(plot_data, paste0("GAM Plot for ", spline_var, " in the model ", vaccine_variable), spline_var)
  print(plot)

  # Guardar el gráfico
  ggsave(
    filename = paste0("~/projects/vacunas_brands_2025/figures/GAM_plot/GAM_", vaccine_variable, "_2spline_", spline_var, ".png"),
    plot = plot,
    width = 8,
    height = 6,
    dpi = 300
  )

  time_elapsed <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  print(paste("Time to run:", time_elapsed, "minutes"))
}

```

## GAM plot wth breakpoints

```{r}
# ============================================================================
# PIECEWISE LINEAR SEGMENTATION
# ============================================================================
segment_curve <- function(x, y, 
                         min_points = 5, 
                         span_val = 0.08, 
                         min_slope_ratio = 2.0,
                         min_prob_change = 5,
                         min_delta_y_breakpoints = 3,
                         min_segment_length_pct = 0.05,
                         compare_bidirectional = TRUE,
                         show_plot = TRUE) {
  library(ggplot2)
  
  # ---- 0. Input validation ----
  if (length(x) != length(y)) {
    stop("x and y must have the same length")
  }
  if (length(x) < min_points * 2) {
    warning("Data too short for segmentation, returning single segment")
    return(list(
      slopes = NA,
      change_points = integer(0),
      segments = list(data.frame(x = x, y = y)),
      n_segments = 1,
      significance_scores = numeric(0),
      filtering_info = list(
        original_breakpoints = 0,
        filtered_breakpoints = 0,
        min_slope_ratio = min_slope_ratio,
        min_prob_change = min_prob_change,
        min_delta_y_breakpoints = min_delta_y_breakpoints,
        kept_most_significant = FALSE
      )
    ))
  }
  
  # ---- 1. LOESS smoothing ----
  loess_fit <- loess(y ~ x, span = span_val)
  y_smooth <- predict(loess_fit)
  
  # ---- 2. Calculate derivative ----
  dy <- diff(y_smooth)
  dx <- diff(x)
  slope <- dy / dx
  
  # ---- 3. Detect local maxima/minima ----
  sign_change <- diff(sign(slope))
  min_indices <- which(sign_change > 0) + 1
  max_indices <- which(sign_change < 0) + 1
  ext_indices <- sort(unique(c(min_indices, max_indices)))
  
  # If no extrema, curve is monotonic
  if (length(ext_indices) == 0) {
    cat("\nNo local extrema detected - curve is monotonic.\n")
    return(list(
      slopes = coef(lm(y_smooth ~ x))[2],
      change_points = integer(0),
      segments = list(data.frame(x = x, y = y_smooth)),
      n_segments = 1,
      significance_scores = numeric(0),
      filtering_info = list(
        original_breakpoints = 0,
        filtered_breakpoints = 0,
        min_slope_ratio = min_slope_ratio,
        min_prob_change = min_prob_change,
        min_delta_y_breakpoints = min_delta_y_breakpoints,
        kept_most_significant = FALSE
      )
    ))
  }
  
  # ---- 4. Create initial segments ----
  breaks <- ext_indices
  segments <- lapply(seq_along(breaks), function(i) {
    start <- ifelse(i == 1, 1, breaks[i - 1] + 1)
    end <- ifelse(i == length(breaks), length(x), breaks[i])
    data.frame(x = x[start:end], y = y_smooth[start:end])
  })
  
  # ---- 5. Calculate statistics for each segment ----
  segment_stats <- data.frame(
    seg = seq_along(segments),
    slope = sapply(segments, function(seg) {
      if (nrow(seg) >= min_points) {
        coef(lm(y ~ x, data = seg))[2]
      } else {
        NA
      }
    }),
    len = sapply(segments, nrow),
    y_start = sapply(segments, function(seg) seg$y[1]),
    y_end = sapply(segments, function(seg) seg$y[nrow(seg)]),
    y_change = sapply(segments, function(seg) abs(seg$y[nrow(seg)] - seg$y[1]))
  )
  
  # ---- 6. Calculate significance score for each breakpoint ----
  significance_scores <- rep(0, length(breaks))
  
  for (i in seq_along(breaks)) {
    # Compare with previous segment (if exists)
    score_prev <- 0
    if (i > 1 && !is.na(segment_stats$slope[i]) && !is.na(segment_stats$slope[i-1])) {
      slope_prev <- abs(segment_stats$slope[i - 1])
      slope_curr <- abs(segment_stats$slope[i])
      if (slope_prev < 1e-6) slope_prev <- 1e-6
      if (slope_curr < 1e-6) slope_curr <- 1e-6
      slope_ratio_prev <- max(slope_curr / slope_prev, slope_prev / slope_curr)
      
      y_change_prev <- segment_stats$y_change[i-1] + segment_stats$y_change[i]
      sign_change_prev <- ifelse(sign(segment_stats$slope[i]) != sign(segment_stats$slope[i-1]), 2, 1)
      
      score_prev <- (slope_ratio_prev * y_change_prev * sign_change_prev) / 10
    }
    
    # Compare with next segment (if exists)
    score_next <- 0
    if (i < nrow(segment_stats) && !is.na(segment_stats$slope[i]) && !is.na(segment_stats$slope[i+1])) {
      slope_curr <- abs(segment_stats$slope[i])
      slope_next <- abs(segment_stats$slope[i + 1])
      if (slope_curr < 1e-6) slope_curr <- 1e-6
      if (slope_next < 1e-6) slope_next <- 1e-6
      slope_ratio_next <- max(slope_next / slope_curr, slope_curr / slope_next)
      
      y_change_next <- segment_stats$y_change[i] + segment_stats$y_change[i+1]
      sign_change_next <- ifelse(sign(segment_stats$slope[i]) != sign(segment_stats$slope[i+1]), 2, 1)
      
      score_next <- (slope_ratio_next * y_change_next * sign_change_next) / 10
    }
    
    # Final score is the maximum between comparisons
    significance_scores[i] <- max(score_prev, score_next)
  }
  
  # ---- 7. Bidirectional filtering ----
  min_segment_length <- length(x) * min_segment_length_pct
  keep_seg <- rep(TRUE, length(breaks))
  
  # First pass: remove very short segments or small changes
  for (i in seq_along(breaks)) {
    seg_idx <- i
    if (seg_idx <= length(segment_stats$slope)) {
      if (is.na(segment_stats$slope[seg_idx]) || 
          segment_stats$len[seg_idx] < min_points ||
          segment_stats$y_change[seg_idx] < min_prob_change) {
        keep_seg[i] <- FALSE
      }
    }
  }
  
  # Second pass: compare with previous and next segments
  for (i in seq_along(breaks)) {
    if (!keep_seg[i]) next
    
    meets_criteria_with_prev <- TRUE
    meets_criteria_with_next <- TRUE
    
    # Comparison with previous segment
    if (i > 1 && keep_seg[i-1]) {
      seg_curr <- i
      seg_prev <- i - 1
      
      if (seg_curr <= length(segment_stats$slope) && 
          seg_prev <= length(segment_stats$slope) &&
          !is.na(segment_stats$slope[seg_curr]) && 
          !is.na(segment_stats$slope[seg_prev])) {
        
        slope_prev <- abs(segment_stats$slope[seg_prev])
        slope_curr <- abs(segment_stats$slope[seg_curr])
        
        if (slope_prev < 1e-6) slope_prev <- 1e-6
        if (slope_curr < 1e-6) slope_curr <- 1e-6
        
        slope_ratio <- max(slope_curr / slope_prev, slope_prev / slope_curr)
        sign_change <- sign(segment_stats$slope[seg_curr]) != sign(segment_stats$slope[seg_prev])
        
        if (slope_ratio < min_slope_ratio && 
            !sign_change && 
            segment_stats$len[seg_curr] < min_segment_length) {
          meets_criteria_with_prev <- FALSE
        }
      }
    }
    
    # Comparison with next segment
    if (i < length(breaks) && keep_seg[i+1]) {
      seg_curr <- i
      seg_next <- i + 1
      
      if (seg_curr <= length(segment_stats$slope) && 
          seg_next <= length(segment_stats$slope) &&
          !is.na(segment_stats$slope[seg_curr]) && 
          !is.na(segment_stats$slope[seg_next])) {
        
        slope_curr <- abs(segment_stats$slope[seg_curr])
        slope_next <- abs(segment_stats$slope[seg_next])
        
        if (slope_curr < 1e-6) slope_curr <- 1e-6
        if (slope_next < 1e-6) slope_next <- 1e-6
        
        slope_ratio <- max(slope_next / slope_curr, slope_curr / slope_next)
        sign_change <- sign(segment_stats$slope[seg_curr]) != sign(segment_stats$slope[seg_next])
        
        if (slope_ratio < min_slope_ratio && 
            !sign_change && 
            segment_stats$len[seg_curr] < min_segment_length) {
          meets_criteria_with_next <- FALSE
        }
      }
    }
    
    # Keep breakpoint if meets criteria with either neighbor
    if (compare_bidirectional) {
      if (!meets_criteria_with_prev && !meets_criteria_with_next) {
        keep_seg[i] <- FALSE
      }
    } else {
      if (!meets_criteria_with_prev) {
        keep_seg[i] <- FALSE
      }
    }
  }
  
  # Third pass: remove consecutive extrema close in Y
  i <- 1
  while (i < length(breaks)) {
    if (!keep_seg[i] || !keep_seg[i+1]) {
      i <- i + 1
      next
    }
    
    y_diff <- abs(y_smooth[breaks[i+1]] - y_smooth[breaks[i]])
    
    if (y_diff < min_prob_change) {
      if (significance_scores[i] < significance_scores[i + 1]) {
        keep_seg[i] <- FALSE
      } else {
        keep_seg[i+1] <- FALSE
      }
    }
    
    i <- i + 1
  }
  
  # Fourth pass: verify delta Y between consecutive breakpoints
  kept_indices <- which(keep_seg)
  
  if (length(kept_indices) > 1) {
    i <- 1
    while (i < length(kept_indices)) {
      idx1 <- breaks[kept_indices[i]]
      idx2 <- breaks[kept_indices[i + 1]]
      
      delta_y <- abs(y_smooth[idx2] - y_smooth[idx1])
      
      if (delta_y < min_delta_y_breakpoints) {
        if (significance_scores[kept_indices[i]] < significance_scores[kept_indices[i + 1]]) {
          keep_seg[kept_indices[i]] <- FALSE
          kept_indices <- which(keep_seg)
        } else {
          keep_seg[kept_indices[i + 1]] <- FALSE
          kept_indices <- which(keep_seg)
        }
      } else {
        i <- i + 1
      }
    }
  }
  
  # ---- 8. If no points remain, keep most significant ----
  kept_indices <- which(keep_seg)
  
  if (length(kept_indices) == 0 && length(breaks) > 0) {
    most_significant_idx <- which.max(significance_scores)
    kept_indices <- most_significant_idx
    
    cat("\n")
    cat("======================================================================\n")
    cat("  WARNING: No breakpoints met all filtering criteria.\n")
    cat("  Keeping the most significant breakpoint (index:", most_significant_idx, ")\n")
    cat("  Significance score:", round(significance_scores[most_significant_idx], 2), "\n")
    cat("======================================================================\n\n")
  }
  
  # ---- 9. Calculate final change points ----
  if (length(kept_indices) > 0) {
    change_points <- breaks[kept_indices]
    change_points <- change_points[!is.na(change_points)]
    change_points <- change_points[change_points > 0 & change_points < length(x)]
    change_points <- sort(unique(change_points))
  } else {
    change_points <- integer(0)
  }
  
  # ---- 10. Create final segments ----
  segments_opt <- list()
  if (length(change_points) == 0) {
    segments_opt[[1]] <- data.frame(x = x, y = y_smooth)
  } else {
    if (change_points[1] > 1) {
      segments_opt[[1]] <- data.frame(
        x = x[1:change_points[1]], 
        y = y_smooth[1:change_points[1]]
      )
    }
    
    if (length(change_points) > 1) {
      for (i in 2:length(change_points)) {
        start_idx <- change_points[i - 1] + 1
        end_idx <- change_points[i]
        
        if (start_idx <= end_idx && start_idx <= length(x) && end_idx <= length(x)) {
          segments_opt[[i]] <- data.frame(
            x = x[start_idx:end_idx],
            y = y_smooth[start_idx:end_idx]
          )
        }
      }
    }
    
    last_start <- change_points[length(change_points)] + 1
    if (last_start <= length(x)) {
      segments_opt[[length(change_points) + 1]] <- data.frame(
        x = x[last_start:length(x)],
        y = y_smooth[last_start:length(y_smooth)]
      )
    }
  }
  
  segments_opt <- segments_opt[sapply(segments_opt, function(s) nrow(s) > 0)]
  
  # ---- 11. Calculate final slopes ----
  slopes_opt <- sapply(segments_opt, function(seg) {
    if(nrow(seg) > 1) {
      coef(lm(y ~ x, data = seg))[2]
    } else {
      NA
    }
  })
  
  # ---- 12. Return results ----
  return(list(
    slopes = slopes_opt,
    change_points = change_points,
    segments = segments_opt,
    n_segments = length(segments_opt),
    significance_scores = significance_scores,
    filtering_info = list(
      original_breakpoints = length(breaks),
      filtered_breakpoints = length(change_points),
      min_slope_ratio = min_slope_ratio,
      min_prob_change = min_prob_change,
      min_delta_y_breakpoints = min_delta_y_breakpoints,
      compare_bidirectional = compare_bidirectional,
      kept_most_significant = length(kept_indices) == 1 && sum(keep_seg, na.rm = TRUE) == 0
    )
  ))
}

# ============================================================================
# GAM FUNCTIONS
# ============================================================================

run_gam <- function(outcome, data, smoothing_term, predictors, vaccine_variable) {
  print(Sys.time())
  predictors <- setdiff(predictors, smoothing_term)
  formula_str <- paste(outcome, " ~ s(", smoothing_term, ") +", vaccine_variable, 
                       " + ", paste(predictors, collapse = " + "))
  formula <- as.formula(formula_str)
  gam_model <- gam(formula, 
                   data = data, 
                   family = binomial(), 
                   method = "REML")
  print(summary(gam_model))
  print(paste("k value used for spline is:", gam_model$smooth[[1]]$bs.dim))
  return(gam_model)
}

extract_gam_data <- function(model, smoothing_term) {
  gam_plot_data <- plot(model, pages = 1)[[1]] 
  x <- gam_plot_data$x
  fit <- gam_plot_data$fit
  se <- gam_plot_data$se
  prob <- exp(fit) / (1 + exp(fit)) * 100 
  lower <- exp(fit - 2 * se) / (1 + exp(fit - 2 * se)) * 100
  upper <- exp(fit + 2 * se) / (1 + exp(fit + 2 * se)) * 100
  data.frame(smoothing_term = x, probability = prob, lower = lower, upper = upper) 
}

create_gam_plot <- function(data, title, smoothing_term) {
  ggplot(data, aes(x = smoothing_term, y = probability)) + 
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
    geom_line(color = "blue") +
    labs(title = title, x = smoothing_term, y = "Probability (%)") +
    ylim(0, 100) + 
    scale_x_continuous(breaks = scales::breaks_width(50)) +
    theme_minimal()
}

create_gam_plot_with_breakpoints_and_histogram <- function(plot_data, 
                                                            raw_data, 
                                                            breakpoints, 
                                                            title, 
                                                            smoothing_term,
                                                            histogram_bins = 30,
                                                            show_breakpoints = TRUE) {
  # Calculate histogram data
  x_raw <- raw_data[[smoothing_term]]
  x_range <- range(plot_data$smoothing_term, na.rm = TRUE)
  
  hist_data <- hist(x_raw, breaks = histogram_bins, plot = FALSE)
  
  # Create histogram dataframe (normalized by total)
  hist_df <- data.frame(
    x = hist_data$mids,
    frequency = hist_data$counts / sum(hist_data$counts) * 100
  )
  
  # Find maximum frequency for scaling
  max_freq <- max(hist_df$frequency, na.rm = TRUE)
  
  # Scale histogram to full 0-100 range
  target_max_height <- 100  # Tallest bar reaches top of plot
  hist_df$frequency_scaled <- (hist_df$frequency / max_freq) * target_max_height
  
  # Calculate percentiles (tertiles and quartiles)
  median_val <- median(x_raw, na.rm = TRUE)
  q1_val <- quantile(x_raw, 0.25, na.rm = TRUE)
  q3_val <- quantile(x_raw, 0.75, na.rm = TRUE)
  t1_val <- quantile(x_raw, 1/3, na.rm = TRUE)
  t2_val <- quantile(x_raw, 2/3, na.rm = TRUE)
  p95_val <- quantile(x_raw, 0.95, na.rm = TRUE)
  
  # Create plot
  p <- ggplot(plot_data, aes(x = smoothing_term, y = probability)) + 
    # IQR shaded region at bottom
    annotate("rect", 
             xmin = q1_val, xmax = q3_val, 
             ymin = 0, ymax = 8,
             fill = "darkgreen", alpha = 0.2) +
    # # Tertile 1 line (segmented to avoid text overlap)
    # annotate("segment",
    #          x = t1_val, xend = t1_val,
    #          y = 9, yend = 82,
    #          linetype = "dashed", 
    #          color = "darkgreen", 
    #          linewidth = 0.7,
    #          alpha = 0.5) +
    # Median line (segmented to avoid text overlap)
    annotate("segment",
             x = median_val, xend = median_val,
             y = 9, yend = 92,
             linetype = "dashed", 
             color = "darkgreen", 
             linewidth = 1,
             alpha = 0.7) +
    # # Tertile 2 line (segmented to avoid text overlap)
    # annotate("segment",
    #          x = t2_val, xend = t2_val,
    #          y = 9, yend = 82,
    #          linetype = "dashed", 
    #          color = "darkgreen", 
    #          linewidth = 0.7,
    #          alpha = 0.5) +
    # p95 line (segmented to avoid text overlap)
    annotate("segment",
             x = p95_val, xend = p95_val,
             y = 0, yend = 80,
             linetype = "dashed", 
             color = "darkgreen", 
             linewidth = 0.7,
             alpha = 0.5) +
    # Histogram bars
    geom_col(data = hist_df, 
             aes(x = x, y = frequency_scaled), 
             fill = "gray70", 
             alpha = 0.4,
             width = diff(hist_data$breaks)[1]) +
    # GAM confidence ribbon
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3, fill = "blue") +
    # GAM probability line
    geom_line(aes(y = probability), color = "blue", linewidth = 1) +
    # Breakpoints (only if show_breakpoints = TRUE and breakpoints exist)
    {if (show_breakpoints && length(breakpoints) > 0) {
      breakpoint_data <- data.frame(x = breakpoints)
      breakpoint_data$y <- sapply(breakpoints, function(bp) {
        idx <- which.min(abs(plot_data$smoothing_term - bp))
        plot_data$probability[idx]
      })
      
      list(
        geom_vline(data = data.frame(x = breakpoints),
                   aes(xintercept = x), 
                   linetype = "dashed", 
                   color = "red", 
                   alpha = 0.6,
                   linewidth = 0.7),
        geom_point(data = breakpoint_data, 
                   aes(x = x, y = y), 
                   color = "red", 
                   size = 3, 
                   shape = 19)
      )
    }} +
    # # Text annotations for tertiles and median
    # annotate("text", x = t1_val, y = 80, 
    #          label = paste0("T1: ", format(round(t1_val, 1), big.mark = ",", scientific = FALSE)), 
    #          color = "darkgreen", size = 3, hjust = 1.2, fontface = "bold", alpha = 0.8) +
    annotate("text", x = median_val, y = 96, 
             label = paste0("Median: ", format(round(median_val, 1), big.mark = ",", scientific = FALSE)), 
             color = "darkgreen", size = 3.5, hjust = 0.5, fontface = "bold") +
    # annotate("text", x = t2_val, y = 80, 
    #          label = paste0("T2: ", format(round(t2_val, 1), big.mark = ",", scientific = FALSE)), 
    #          color = "darkgreen", size = 3, hjust = -0.2, fontface = "bold", alpha = 0.8) +
    annotate("text", x = p95_val, y = 80, 
             label = paste0("P95: ", format(round(p95_val, 1), big.mark = ",", scientific = FALSE)), 
             color = "darkgreen", size = 3, hjust = 0.5, fontface = "bold", alpha = 0.8) +
    annotate("text", x = mean(c(q1_val, q3_val)), y = 4, 
             label = paste0("IQR: ", format(round(q1_val, 1), big.mark = ",", scientific = FALSE), 
                           " - ", format(round(q3_val, 1), big.mark = ",", scientific = FALSE)), 
             color = "darkgreen", size = 3, fontface = "bold") +
    # Dual Y axes
    scale_y_continuous(
      name = "Probability (%)",
      limits = c(0, 100),
      sec.axis = sec_axis(
        trans = ~ . * max_freq / target_max_height,
        name = "Frequency (% of sample)",
        breaks = seq(0, ceiling(max_freq), length.out = 6)
      )
    ) + 
    scale_x_continuous(breaks = scales::breaks_width(50)) +
    labs(title = paste(title, "- with Distribution"), 
         x = smoothing_term) +
    theme_minimal() +
    theme(
      axis.title.y.right = element_text(color = "gray50"),
      axis.text.y.right = element_text(color = "gray50")
    )
  
  return(p)
}

# ============================================================================
# MAIN FUNCTION
# ============================================================================

gam_analysis_with_segmentation <- function(data, outcome, smoothing_term, 
                                           predictors, vaccine_variable,
                                           min_points = 5,
                                           min_slope_ratio = 2.0,
                                           min_prob_change = 5,
                                           min_segment_length_pct = 0.05,
                                           histogram_bins = 30,
                                           show_breakpoints = TRUE) { 
  start_time <- Sys.time()
  print(paste("Analysis of", smoothing_term))
  
  # 1. Run GAM model
  print(paste("Running GAM model for:", outcome))
  model <- run_gam(outcome, data, smoothing_term, predictors, vaccine_variable) 
  plot_data <- extract_gam_data(model, smoothing_term) 
  plot_data$outcome <- outcome 
  
  # 2. Apply piecewise linear segmentation
  print("Applying piecewise linear segmentation...")
  seg_result <- segment_curve(
    x = plot_data$smoothing_term,
    y = plot_data$probability,
    min_points = min_points,
    min_slope_ratio = min_slope_ratio,
    min_prob_change = min_prob_change,
    min_segment_length_pct = min_segment_length_pct
  )
  
  # 3. Display breakpoint table
  cat("\n")
  cat("======================================================================\n")
  cat("  BREAKPOINTS ON X-AXIS (Segmentation)\n")
  cat("======================================================================\n\n")
  
  if (length(seg_result$change_points) > 0) {
    break_x <- plot_data$smoothing_term[seg_result$change_points]
    break_prob <- plot_data$probability[seg_result$change_points]
    
    breaks_table <- data.frame(
      Segment = 1:length(break_x),
      Breakpoint_X = round(break_x, 2),
      Probability_Y = round(break_prob, 2)
    )
    names(breaks_table) <- c("Segment", 
                             paste("Breakpoint at", smoothing_term),
                             "Probability (%)")
    print(breaks_table, row.names = FALSE)
    
    if (seg_result$filtering_info$kept_most_significant) {
      cat("\n  Note: Only the most significant breakpoint was kept.\n")
    }
  } else {
    cat("  No breakpoints detected (curve is monotonic).\n")
  }
  
  cat("\n======================================================================\n\n")
  
  # 4. Display filtering summary
  cat("======================================================================\n")
  cat("  FILTERING SUMMARY\n")
  cat("======================================================================\n")
  cat("  Original breakpoints detected:", seg_result$filtering_info$original_breakpoints, "\n")
  cat("  Breakpoints after filtering:", seg_result$filtering_info$filtered_breakpoints, "\n")
  cat("  Number of segments:", seg_result$n_segments, "\n")
  cat("  Parameters used:\n")
  cat("    - min_slope_ratio:", min_slope_ratio, "\n")
  cat("    - min_prob_change:", min_prob_change, "%\n")
  cat("    - min_segment_length_pct:", min_segment_length_pct * 100, "%\n")
  cat("======================================================================\n\n")
  
  # 5. Create and print basic plot
  gam_plot <- create_gam_plot(plot_data, outcome, smoothing_term)
  print(gam_plot)
  
  # 6. Create and save plot with histogram (and optionally breakpoints)
  if (length(seg_result$change_points) > 0) {
    break_x <- plot_data$smoothing_term[seg_result$change_points]
  } else {
    break_x <- numeric(0)
  }
  
  gam_plot_with_histogram <- create_gam_plot_with_breakpoints_and_histogram(
    plot_data = plot_data,
    raw_data = data,
    breakpoints = break_x,
    title = outcome, 
    smoothing_term = smoothing_term,
    histogram_bins = histogram_bins,
    show_breakpoints = show_breakpoints
  )
  print(gam_plot_with_histogram)
  
  # Adjust filename based on whether breakpoints are shown
  if (show_breakpoints) {
    filename_suffix <- "_breakpoints"
  } else {
    filename_suffix <- "_distribution"
  }
  
  ggsave(gam_plot_with_histogram,
         filename = paste0("~/projects/vacunas_brands_2025/figures/GAM_plot/GAM_", 
                           smoothing_term, "_", vaccine_variable, ".png"),
         device = "jpeg", units = "px", dpi = 300, width = 3000, height = 2000)
  
  # 7. Print elapsed time
  time_elapsed <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  print(paste("Time to run:", time_elapsed, "minutes"))
  print(Sys.time())
  
  # 8. Return results
  return(list(
    plot = gam_plot,
    plot_with_histogram = gam_plot_with_histogram,
    model = model,
    segmentation = seg_result,
    plot_data = plot_data
  ))
}

```

## GAM plot - interactive

```{r}
# --- Funciones auxiliares ---
extract_gam_data_interactive <- function(model, spline_var) {
  gp <- plot(model, term = spline_var, pages = 1)[[1]]
  x   <- gp$x
  fit <- gp$fit
  se  <- gp$se

  prob  <- plogis(fit) * 100
  lower <- plogis(fit - 2*se) * 100
  upper <- plogis(fit + 2*se) * 100

  tibble(x = x, probability = prob, lower = lower, upper = upper)
}

create_gam_plotly <- function(df, title, xlab) {
  # índices de mínimo y máximo
  idx_min <- which.min(df$probability)
  idx_max <- which.max(df$probability)
  ext_df  <- df[c(idx_min, idx_max), ] %>%
    mutate(extreme = c("Min", "Max"))

  p <- ggplot(df, aes(x = x, y = probability)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.3) +
    geom_line() +
    geom_point(data = ext_df, size = 2) +
    geom_vline(xintercept = ext_df$x, linetype = "dotted") +
    labs(
      title = title,
      x = xlab,
      y = "Probability (%)",
      caption = "*Comorbidities, sex and SARS‑CoV‑2 variant period were also included as confounders"
    ) +
    ylim(0, 100) +
    theme_minimal() +
    theme(plot.caption = element_text(hjust = 0, face = "italic", size = 9))

  # Tooltips personalizados
  ply <- ggplotly(
    p,
    tooltip = "text"
  ) %>%
    style(
      text = paste0(
        xlab, ": ", round(df$x, 1),
        "<br>Probability: ", round(df$probability, 1), "%",
        "<br>95% CI: [", round(df$lower, 1), ", ", round(df$upper, 1), "]"
      ),
      traces = 2 # traza de la línea
    ) %>%
    style(
      text = paste0(
        ext_df$extreme, "<br>",
        xlab, ": ", round(ext_df$x, 1),
        "<br>Probability: ", round(ext_df$probability, 1), "%"
      ),
      traces = 3 # traza de los puntos extremo
    )

  return(ply)
}

# --- Función principal ---
generate_gam_plot_interactive <- function(data, vaccine_variable, spline_var) {
  message("GAM of Clinical Event")
  start_time <- Sys.time(); print(start_time)

  if (spline_var == "age") {
    order_smooth <- "s(age) + s(ultima_inmun_infecc_dias)"
    xlab <- "Age"
  } else if (spline_var == "ultima_inmun_infecc_dias") {
    order_smooth <- "s(ultima_inmun_infecc_dias) + s(age)"
    xlab <- "Days since last immunization/infection"
  } else {
    stop("spline_var debe ser 'age' o 'ultima_inmun_infecc_dias'")
  }

  formula_text <- paste0(
    "clinical_event ~ ", order_smooth, " + ",
    vaccine_variable, " + desc_sexo + variant_period + asma + cardiopatia_isquemica + ",
    "cirrosis_hepatica + diabetes + epoc + hepatopatia_cronica_excepto_cirrosis + hipertension + ",
    "insuficiencia_cardiaca + insuficiencia_renal_cronica + vih + cancer_hematologico + cancer_organo_solido"
  )
  model <- gam(as.formula(formula_text), data = data, family = binomial(), method = "REML")
  print(summary(model))

  plot_data <- extract_gam_data_interactive(model, spline_var)
  ply <- create_gam_plotly(
    plot_data,
    paste0("GAM (", spline_var, "). Probability vs ", xlab, " — model: ", vaccine_variable),
    xlab
  )

  # Guardar como HTML interactivo
  out_path <- paste0("~/projects/vacunas_brands_2025/figures/GAM_plot/GAM_", vaccine_variable, "_2spline_", spline_var, ".html")
  saveWidget(ply, file = out_path, selfcontained = TRUE)

  time_elapsed <- round(difftime(Sys.time(), start_time, units = "mins"), 2)
  message("Time to run: ", time_elapsed, " minutes")
  return(ply)
}
```

# Forest plot - GAM - 1 spline (age)

## Reduced variables in forest plot
### vac_1shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_1shot",
  variables = "reduced",
  file_name = "vac1shot_reduced",
  title = "One-dose COVID-19 vaccination regimen"
  )

```

### vac_2shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_2shot",
  variables = "reduced",
  file_name = "vac2shot_reduced",
  title = "Two-dose COVID-19 vaccination regimen"
  )

```

### vac_3shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_3shot",
  variables = "reduced",
  file_name = "vac3shot_reduced",  
  title = "Three-dose COVID-19 vaccination regimen"
  )

```
## All variables in forest plot

### vac_1shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_1shot",
  variables = "all",
  file_name = "vac1shot_all",
  title = "One-dose COVID-19 vaccination regimen"
  )

```

### vac_2shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_2shot",
  variables = "all",
  file_name = "vac2shot_all",
  title = "Two-dose COVID-19 vaccination regimen"
  )

```

### vac_3shot
```{r}
log_reg_gam_1s(
  data = final_dataset, 
  vaccine_variable = "vac_3shot",
  variables = "all",
  file_name = "vac3shot_all",  
  title = "Three-dose COVID-19 vaccination regimen"
  )

```

# Forest plot - GAM - 2 splines (not used)

## vac_1shot
```{r}
log_reg_gam_2s(
  data = final_dataset, 
  vaccine_variable = "vac_1shot", 
  file_name = "vac1shot",
  title = "1-shot Regimen - Splines for Age and Last Immunization", 
  number_splines = 2)

```

## vac_2shot
```{r}
log_reg_gam_2s(
  data = final_dataset, 
  vaccine_variable = "vac_2shot",
  file_name = "vac2shot",
  title = "2-shot Regimen - Splines for Age and Last Immunization", 
  number_splines = 2)

```

## vac_3shot
```{r}
log_reg_gam_2s(
  data = final_dataset, 
  vaccine_variable = "vac_3shot",
  file_name = "vac3shot",  
  title = "3-shot Regimen - Splines for Age and Last Immunization", 
  number_splines = 2)

```


# GAM Plot of last immunization

## vac_1shot - breakpoints

```{r}
# Call the GAM function without breakpoints
result_with_breakpoints <- gam_analysis_with_segmentation(
  data = final_dataset,
  outcome = "clinical_event",
  smoothing_term = "ultima_inmun_infecc_dias",
  predictors = predictor_vars,
  vaccine_variable = "vac_1shot", 
  min_points = 0,
  min_slope_ratio = 0,
  min_prob_change = 0,
  min_segment_length_pct = 0.0,
  histogram_bins = 100,
  show_breakpoints = TRUE  # True for showing red breakpoint lines and points
)

```

## vac_2shot - breakpoints

```{r}
# Call the GAM function without breakpoints
result_with_breakpoints <- gam_analysis_with_segmentation(
  data = final_dataset,
  outcome = "clinical_event",
  smoothing_term = "ultima_inmun_infecc_dias",
  predictors = predictor_vars,
  vaccine_variable = "vac_2shot", 
  min_points = 0,
  min_slope_ratio = 0,
  min_prob_change = 0,
  min_segment_length_pct = 0.0,
  histogram_bins = 100,
  show_breakpoints = TRUE  # True for showing red breakpoint lines and points
)

```

## vac_3shot - breakpoints

```{r}
# Call the GAM function without breakpoints
result_with_breakpoints <- gam_analysis_with_segmentation(
  data = final_dataset,
  outcome = "clinical_event",
  smoothing_term = "ultima_inmun_infecc_dias",
  predictors = predictor_vars,
  vaccine_variable = "vac_3shot", 
  min_points = 0,
  min_slope_ratio = 0,
  min_prob_change = 0,
  min_segment_length_pct = 0.0,
  histogram_bins = 100,
  show_breakpoints = TRUE  # True for showing red breakpoint lines and points
)

```

## vac_1shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_1shot",
                  spline_var = "ultima_inmun_infecc_dias")



```

## vac_2shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_2shot",
                  spline_var = "ultima_inmun_infecc_dias")
```


## vac_3shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_3shot",
                  spline_var = "ultima_inmun_infecc_dias")
```

# GAM Plot of age

## vac_1shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_1shot",
                  spline_var = "age")
```

## vac_2shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_2shot",
                  spline_var = "age")
```


## vac_3shot
```{r}
generate_gam_plot(data = final_dataset,
                  vaccine_variable = "vac_3shot",
                  spline_var = "age")
```

